Document Embedding for Scientific Articles:\newline A validation of word embeddings

### 2018-07-25 Paper/paper.tex
overleaf sync

\institute{University of Amsterdam, Science park 904, 1012WX Amsterdam, The Netherlands \and
Elsevier, Radarweg 29, 1043 NX Amsterdam, The Neterlands
\email{meijerarjan@live.nl, r.karimi@elsevier.com}}
\maketitle

@@

For this research we used a dataset consisting of $186.962.354$ tokens. These tokens have been collected from $1.391.543$ articles which have been published in $3.759$ distinct journals. In our dataset, each article is represented with a title and an abstract. The word occurrences in our dataset follow a Pareto-like distribution as described by \citet{wiegand2018word}, indicating that our original data has the same properties as standard English texts. We used Spark for this research, enabeling us to use its tools for big data processing\cite{armbrust2015spark}.

@@

We used 300-dimensional embeddings, created by \citet{Truong2017Thesis}, based on the Word2Vec model. From these embeddings we created multiple embedding sets. These sets are: the default embedding; as created by \citet{Truong2017Thesis} and TF-IDF weighted embeddings (referred to in figures as \textit{embedding}). We used 4 variants of TF-IDF weighted embeddings: 

@@

The TF-IDF feature vectors outperform the embeddings on the abstracts, while the embeddings outperform the TF-IDF feature vectors on the titles. The main difference between the abstract and title is that the title contains fewer tokens compared to the abstract (see Table~\ref{table:corpusSize}). This means that the titles contain less information than the abstracts. Due to this, the TF-IDF method, which is purely based on word-occurrences \& counts has less information on the titles. The TF-IDF method works better on the abstract, containing more tokens than the tile, thus improving the differentiating the different abstract. The data furthermore shows that increasing the vocabulary size increases the performance of the TF-IDF, meaning that none of the created cut-off's resulted in cutting off noise, the increasing size of the vocabulary only improved the performance on our data. It could be possible that at higher vocabulary sizes the cut-off would result in a sharper signal, but we did not investigate this further due to our findings that the TF-IDF performance stagnates (presented in Figure~\ref{figure:tfidfPerformance}).

@@

This research shows that even though the embeddings can capture and preserve relatedness, TF-IDF is able to outperform the embeddings on the abstracts. Earlier research already proposed improvements to the word embeddings. \citet{dai2015document} show that using paragraph vectors improves the accuracy of word embeddings with 4.4\% on triplet creation with the Wikipedia corpus and a 3.9\% improvement on the same task based on the arXiv articles.\\
Furthermore, \citet{le2014distributed} show that the usage of paragraph vectors decrease the error rate (positive/negative) with 7.7\% compared to averaging the word embeddings on categorizing text as either positive or negative. While the improvement looks promising, we have to keep in mind that our task differs from earlier tasks. We do not categorize on two categories but more than 3k. Still, we would expect an improvement by using paragraph vectors since the classification task if fundamentally the same, only on a much larger scale, which complicates the task due to the "grey areas"  between categories. These are the area's in which the classification algorithm is "in doubt" and could reasonably assign the article to both journals. The number of these grey areas increase given more categories. \citet{pennington2014glove} showed that the GloVe model outperforms the CBOW model, which is used in this research, on a word analogy task. \citet{wang2016linked} introduced the Linked Document Embedding method (LDE) method, which makes use of additional information about a document, such as citations. Their research specifically focused on categorizing documents, showed a 5.89\% increase of the micro-F1 score on LDE compared to CBOW, and a 9.11\% increase of the macro-F1 score. We would expect that applying this technique to our dataset would improve our scores, given earlier results on comparable tasks.\\

@@

%\begin{theorem}
%This is a sample theorem. The run-in heading is set in bold, while
%the following text appears in italics. Definitions, lemmas,
%propositions, and corollaries are styled the same way.
%\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
%\begin{proof}
%Proofs, examples, and remarks have the initial word in italics,
%while the following text appears in normal font.
%\end{proof}

@@

\bibliography{Bibliography}

>>>>

\institute{University of Amsterdam, Science park 904, 1012WX Amsterdam, The Netherlands \and
Elsevier, Radarweg 29, 1043 NX Amsterdam, The Neterlands
\email{meijerarjan@live.nl, r.karimi@elsevier.com}}
\maketitle

@@

For this research we used a dataset consisting of $186.962.354$ tokens. These tokens have been collected from $1.391.543$ articles which have been published in $3.759$ distinct journals. In our dataset, each article is represented with a title and an abstract. The word occurrences in our dataset follow a Pareto-like distribution as described by Wiegand et al\cite{wiegand2018word}, indicating that our original data has the same properties as standard English texts. We used Spark for this research, enabeling us to use its tools for big data processing\cite{armbrust2015spark}.

@@

We used 300-dimensional embeddings, created by Truong\cite{Truong2017Thesis}, based on the Word2Vec model. From these embeddings we created multiple embedding sets. These sets are: the default embedding; as created by Truong\cite{Truong2017Thesis} and TF-IDF weighted embeddings (referred to in figures as \textit{embedding}). We used 4 variants of TF-IDF weighted embeddings: 

@@

The TF-IDF feature vectors outperform the embeddings on the abstracts, while the embeddings outperform the TF-IDF feature vectors on the titles. The main difference between the abstract and title is that the title contains fewer tokens compared to the abstract. This means that the titles contain less information than the abstracts. Due to this, the TF-IDF method, which is purely based on word-occurrences \& counts has less information on the titles. The TF-IDF method works better on the abstract, containing more tokens than the tile, thus improving the differentiating the different abstract. The data furthermore shows that increasing the vocabulary size increases the performance of the TF-IDF, meaning that none of the created cut-off's resulted in cutting off noise, the increasing size of the vocabulary only improved the performance on our data. It could be possible that at higher vocabulary sizes the cut-off would result in a sharper signal, but we did not investigate this further due to our findings that the TF-IDF performance stagnates (presented in Figure~\ref{figure:tfidfPerformance}).

@@

This research shows that even though the embeddings can capture and preserve relatedness, TF-IDF is able to outperform the embeddings on the abstracts. Earlier research already proposed improvements to the word embeddings. Dai et al\cite{dai2015document} show that using paragraph vectors improves the accuracy of word embeddings with 4.4\% on triplet creation with the Wikipedia corpus and a 3.9\% improvement on the same task based on the arXiv articles.\\
Furthermore, Le et al\cite{le2014distributed} show that the usage of paragraph vectors decrease the error rate (positive/negative) with 7.7\% compared to averaging the word embeddings on categorizing text as either positive or negative. While the improvement looks promising, we have to keep in mind that our task differs from earlier tasks. We do not categorize on two categories but more than 3k. Still, we would expect an improvement by using paragraph vectors since the classification task if fundamentally the same, only on a much larger scale, which complicates the task due to the "grey areas"  between categories. These are the area's in which the classification algorithm is "in doubt" and could reasonably assign the article to both journals. The number of these grey areas increase given more categories. Pennington et al\cite{pennington2014glove} showed that the GloVe model outperforms the CBOW model, which is used in this research, on a word analogy task. Wang et al\cite{wang2016linked} introduced the Linked Document Embedding method (LDE) method, which makes use of additional information about a document, such as citations. Their research specifically focused on categorizing documents, showed a 5.89\% increase of the micro-F1 score on LDE compared to CBOW, and a 9.11\% increase of the macro-F1 score. We would expect that applying this technique to our dataset would improve our scores, given earlier results on comparable tasks.\\

@@

@@

\bibliography{references}
